detect patches

### 2024-01-26 scripts/classify_patches.Rmd
rescaling patch detection to data size

```{r}
p <- '../results/suitability_maps'
r <- rast(file.path(p, 'Acamptopappus_sphaerocephalus1k-2024-01-10_08:57:46.tif'))
msk <- ifel(r < 0.80, NA, r)
m <- mask(r, msk)

m <- mask(m, crop(hydr_bound, m), inverse = TRUE)
plot(m)
m1 <- aggregate(m, 2, fun = 'mean')

# aggregate the raster to speed up the patch detection process. 
```

aggregation by a factor of 2 allows for quick patch delineation and detection.

We need to identify and names patches.

```{r}
system.time({ob <- terra::patches(m1,  directions = 4, allowgaps = FALSE)})
system.time({sizes =  zonal(cellSize(ob, unit="ha"), ob, sum, as.raster=TRUE) }) # use this has a mask
size_mask = ifel(sizes < 2.0235, NA, sizes) # remove areas < 5 acres.
ob <- mask(ob, z)
plot(ob)
```

@@

p <- '../results/suitability_maps'
f <- file.path(p, list.files(p, pattern = '1k-'))

@@

  # read in file
  r <- rast(x)
  taxon <- gsub('1k.*$', '', basename(x))
  pout <- '../results/patches'

@@

  # detect patches
  r1 <- terra::aggregate(r, 2, fun = 'mean')
  pat <- terra::patches(r1,  directions = 4, allowGaps = FALSE)

@@

lapply(f[4:335], patcheR)

>>>>

@@

p <- '../results/suitability_maps'
f <- file.path(p, list.files(p, pattern = '1k-'))
info <- lapply(f, file.info)

info1 <- info[[1]]$size/1e9

@@

  # read in file
  r <- rast(x)
  taxon <- gsub('1k.*$', '', basename(x))
  pout <- '../results/patches'
  fsize <- file.info(f[1])$size/1e9

@@

  # detect patches - this scales to the extent of analysis. #
  # Areas need to be aggregateed for computatations to be run in meaningful amount of
  # time

  if(fsize < 0.2){r1 <- r} else
    if(fsize < 0.4){r1 <- terra::aggregate(r, 2, fun = 'mean')} else 
      if(fsize < 0.7){r1 <- terra::aggregate(r, 3, fun = 'mean')} else
        {r1 <- terra::aggregate(r, 4, fun = 'mean')}

  pat <- terra::patches(r1,  directions = 4, allowGaps = FALSE)

@@

lapply(f, patcheR)
